---
title: "Data Processing"
output: 
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
All data processing happens here: Products are saved in Processed data folder. All drought response data is based on BAI growth and lower 10th percentile spei years, but this can be changed here.

### Products
* detrended tree rings
* annual climate data including SPEI
* monthly climate data including SPEI
* tree ring data joined with field data (diameter, competition etc.)
* RRR framework data
* Drought effect framework data
* Recovery (legacy) data

```{r packages, message=FALSE, warning=FALSE}
library(dplR)
library(tidyverse)
library(kableExtra)
library(knitr)
library(SPEI)
library(googledrive)
library(googlesheets4)
library(brms)

```
## Site Characteristics
For consistent naming across data files. Note I'm also reading in the field data here to be manipulated later to keep the google code together

```{r google, echo=T, results='hide'}
## do authentication step first and copy id from google sheet url.
drive_auth(cache = ".secrets")
options(
  gargle_oauth_cache = ".secrets",
  gargle_oauth_email = "tbohn001@ucr.edu"
)

sheets_auth(cache = ".secrets")
options(
  gargle_oauth_cache = ".secrets",
  gargle_oauth_email = "tbohn001@ucr.edu"
)

fielddata <- drive_get(as_id("1Y_PSgFJz4bSW00JYoRiVnXrCrhL776QA4dNSOWl4Clg"))
dia_data <- read_sheet(fielddata, sheet = 2)
sitechar <- read_sheet(fielddata, sheet=3)
dia_data_new <- read_sheet(fielddata, sheet = 4)

sitechar <- rename(sitechar, siteno=No.) %>% 
  dplyr::select(c("Region", "siteno", "Site", "Site2", "Neighborhood", "hilo", "X", "Y"))

sites <- dplyr::select(sitechar, Site2, siteno, Region, hilo) %>% 
  group_by(Site2) %>% 
  summarise_all(first)

sites2 <- dplyr::select(sitechar, Site, Site2, siteno, Region, hilo) %>% 
  group_by(Site2) %>% 
  summarise_all(first)

```

## Tree Ring Data

#### Note on missing rings

When there is a missing ring in one of the two cores growth information is only taken from the core with growth information (not averaged).  When the ring is missing from both core samples the missing growth value is replaced with the mean growth for that tree but denoted as a missing ring for future analyses.

```{r data, echo=T, results='hide'}
bm <- read.rwl("Raw Data/RWL copies/Black_Mountain.rwl")
names(bm)[str_detect(names(bm), "11PL1_3A")==T] <- "11PL1_1C"
names(bm)[str_detect(names(bm), "11PL1_3B")==T] <- "11PL1_1D"

sp <- read.rwl("Raw Data/RWL copies/Seven_Pines.rwl")
bm_ids <- read.ids(bm, stc=c(2,5,1))
sp_ids <- read.ids(sp, stc=c(2,5,1))

bm_id_df <- bm_ids %>%
  mutate(label=rownames(bm_ids),
         Site="bm",
         Species=str_sub(label, 3, 4),
         Neighborhood=str_sub(label, 5, 5),
         ID=ifelse(str_detect(label, "\\_")==TRUE, str_sub(label, 7, 7), str_sub(label, 6,7)),
         tree=as.character(tree))

bm_id_df_tree <- group_by(bm_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

sp_id_df <- sp_ids %>%
  mutate(label=rownames(sp_ids),
         Site="sp",
         Species=str_sub(label, 3, 4),
         Neighborhood=str_sub(label, 5, 5),
         ID=ifelse(str_detect(label, "\\_")==TRUE, str_sub(label, 7, 7), str_sub(label, 6,7)),
         tree=as.character(tree))
sp_id_df_tree <- group_by(sp_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

lc <- read.rwl("Raw Data/RWL copies/Lilly_Creek.rwl")
pp <- read.rwl("Raw Data/RWL copies/Peppermint.rwl")
lc_ids <- read.ids(lc, stc=c(2,5,1))
pp_ids <- read.ids(pp, stc=c(2,5,1))

lc_id_df <- lc_ids %>%
  mutate(label=rownames(lc_ids),
         Site="lc",
         Species=str_sub(label, 3, 4),
         Neighborhood=str_sub(label, 5, 5), 
         ID=str_sub(label, 7, 7),
         tree=as.character(tree))
lc_id_df_tree <- group_by(lc_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

pp_id_df <- pp_ids %>%
  mutate(label=rownames(pp_ids),
         Site="pp",
         Species=str_sub(label, 3, 4),
         Neighborhood=str_sub(label, 5, 5), 
         ID=str_sub(label, 7, 7),
         tree=as.character(tree))
pp_id_df_tree <- group_by(pp_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

cm <- read.rwl("Raw Data/RWL copies/Crescent_Meadow.rwl")
names(cm)[str_detect(names(cm), "4PJ2_1A")==T] <- "4PP2_1A"
names(cm)[str_detect(names(cm), "4PJ2_1B")==T] <- "4PP2_1B"
pr <- read.rwl("Raw Data/RWL copies/Pine_Ridge.rwl")
cm_ids1 <- read.ids(cm[,which(str_length(colnames(cm))==7)], stc=c(1,5,1))
cm_ids2 <- read.ids(cm[,which(str_length(colnames(cm))==8)], stc=c(1,6,1))
cm_ids2$tree <- cm_ids2$tree + max(cm_ids1$tree)
cm_ids <- rbind(cm_ids1, cm_ids2)

pr_ids1 <- read.ids(pr[,which(str_length(colnames(pr))==7)], stc=c(1,5,1))
pr_ids2 <- read.ids(pr[,which(str_length(colnames(pr))==8)], stc=c(1,6,1))
pr_ids2$tree <- pr_ids2$tree + max(pr_ids1$tree)
pr_ids3 <- read.ids(pr[,which(str_length(colnames(pr))==6)], stc=c(1,5,5))
pr_ids3$tree <- pr_ids3$tree + max(pr_ids2$tree)
pr_ids <- rbind(pr_ids1, pr_ids2, pr_ids3)

cm_id_df <- cm_ids %>%
  mutate(label=rownames(cm_ids),
         Site="cm",
         Species=str_sub(label, 2, 3),
         tree=as.character(tree))
cm_id_df$Neighborhood <- NA
for(i in 1:nrow(cm_id_df)) {
  if(str_detect(cm_id_df$label[i], "_")==TRUE) {
    cm_id_df$Neighborhood[i] <- str_sub(cm_id_df$label[i], 4, 4)
  } else NA
}
cm_id_df$ID <- NA
for(i in 1:nrow(cm_id_df)) {
  if(str_detect(cm_id_df$label[i], "_")==TRUE) {
    cm_id_df$ID[i] <- str_sub(cm_id_df$label[i], 6, 6)
  } else cm_id_df$ID[i] <- str_sub(cm_id_df$label[i], 4, 7)
}
cm_id_df$ID <- str_replace_all(cm_id_df$ID, "A", "")
cm_id_df$ID <- str_replace_all(cm_id_df$ID, "B", "")
cm_id_df$Neighborhood <- ifelse(cm_id_df$ID%in%c(102, 105, 117, 118, 122), 1, 
                                ifelse(cm_id_df$ID%in%c(1924, 3064, 3085, 3101), 2,
                                       ifelse(cm_id_df$ID%in%c(1954, 1959, 1965, 1966, 1967), 3, cm_id_df$Neighborhood))) 

cm_id_df_tree <- group_by(cm_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

pr_id_df <- pr_ids %>%
  mutate(label=rownames(pr_ids),
         Site="pr",
         Species=str_sub(label, 2, 3),
         tree=as.character(tree))
pr_id_df$Neighborhood <- NA
for(i in 1:nrow(pr_id_df)) {
  if(str_detect(pr_id_df$label[i], "_")==TRUE) {
    pr_id_df$Neighborhood[i] <- str_sub(pr_id_df$label[i], 4, 4)
  } else NA
}
pr_id_df$ID <- NA
for(i in 1:nrow(pr_id_df)) {
  if(str_detect(pr_id_df$label[i], "_")==TRUE) {
    pr_id_df$ID[i] <- str_sub(pr_id_df$label[i], 6, 6)
  } else pr_id_df$ID[i] <- str_sub(pr_id_df$label[i], 4, 7)
}
pr_id_df$ID <- str_replace_all(pr_id_df$ID, "A", "")
pr_id_df$ID <- str_replace_all(pr_id_df$ID, "B", "")

pr_id_df$ID[which(pr_id_df$ID==5434)] <- 543

pr_id_df$Neighborhood <- ifelse(pr_id_df$ID%in%c(497, 3150, 3157), 1,
                                ifelse(pr_id_df$ID%in%c(528, 533, 543, 552, 3001), 2,
                                       ifelse(pr_id_df$ID%in%c(598, 604), 3, pr_id_df$Neighborhood)))

pr_id_df_tree <- group_by(pr_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))


ic <- read.rwl("Raw Data/RWL copies/Inyo_Craters.rwl")
names(ic)[str_detect(names(ic), "9AC1_1A")==T] <- "9AM1_1A"
names(ic)[str_detect(names(ic), "9AC1_1B")==T] <- "9AM1_1B"
sl <- read.rwl("Raw Data/RWL copies/Sherwin_Lakes.rwl")
ic_ids <- read.ids(ic, stc=c(1,5,1))
sl_ids <- read.ids(sl, stc=c(2,5,1))

ic_id_df <- ic_ids %>%
  mutate(label=rownames(ic_ids),
         Site="ic",
         Species=str_sub(label, 2, 3),
         Neighborhood=str_sub(label, 4, 4), 
         ID=str_sub(label, 6, 6),
         tree=as.character(tree))
ic_id_df_tree <- group_by(ic_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

sl_id_df <- sl_ids %>%
  mutate(label=rownames(sl_ids),
         Site="sl",
         Species=str_sub(label, 3, 4),
         Neighborhood=str_sub(label, 5, 5), 
         ID=str_sub(label, 7, 7),
         tree=as.character(tree))
sl_id_df_tree <- group_by(sl_id_df, tree, Site, Species, Neighborhood, ID) %>%
  summarise(label=first(label))

for(i in c("sp", "bm", "lc", "pp", "pr", "cm", "ic", "sl")){ 
  temp <- get(i)
  temp[temp == 0] <- NA
  treeave <- treeMean(temp, get(paste(i,"_ids", sep="")), na.rm = T)
  zero <- fill.internal.NA(treeave, 0)
  zero$Year <- as.numeric(rownames(zero))

  treeave <- fill.internal.NA(treeave, "Mean")
  raw <- treeave
  raw$Year <- as.numeric(rownames(raw))
  
  bai <- bai.out(raw, diam = NULL)
  bai$Year <- as.numeric(rownames(bai))
  
  spl <- detrend(rwl = treeave, method="Spline")
  spl$Year <- as.numeric(rownames(spl))
  
  fss <- detrend(rwl = treeave, method="Friedman")
  fss$Year <- as.numeric(rownames(fss))
  
  dat<- gather(zero, key="tree", "temp", -Year) %>%
    mutate(zeroRing=ifelse(temp==0, "yes", "no")) %>% 
    dplyr::select(-temp) %>% 
    left_join(gather(raw, key="tree", "raw_growth", -Year)) %>%
    left_join(gather(bai, key="tree", "BAI", -Year)) %>%
    left_join(gather(spl, key="tree", "spline_growth", -Year)) %>%
    left_join(gather(fss, key="tree", "fss_growth", -Year)) %>%
    na.omit() %>%
    left_join(get(paste(i,"_id_df_tree", sep=""))) %>%
    filter(Year>=1895)
  
  if(i=="sp"){allrings_loop <- dat
  } else allrings_loop <- rbind(allrings_loop, dat)
  
}

allrings <- allrings_loop %>% 
  left_join(sites, by=c("Site"="Site2")) %>% 
  mutate(tree.uniqueID = tolower(paste(paste(siteno, Species, Neighborhood, sep=""), ID, sep="_"))) %>% 
  dplyr::select( -tree) %>% 
  mutate_at(c("Neighborhood", "ID"), as.numeric)

write.csv(allrings, "Processed Data/detrended_rings_allsites.csv", row.names=F)
```

## Crossdating summary stats
```{r summarystats, echo=FALSE}
sj_summary <- bind_rows(rwi.stats(sp, sp_ids, prewhiten=T)%>% mutate(site='seven pines'),
                        rwi.stats(bm, bm_ids, prewhiten=T)%>% mutate(site='black mountain'))
sj_summary <- sj_summary[c(14, 1:13)]
kable(sj_summary, caption = "San Jacinto") %>% kable_styling()

senf_summary <- bind_rows(rwi.stats(lc, lc_ids, prewhiten=T)%>% mutate(site='lilly creek'),
                        rwi.stats(pp, pp_ids, prewhiten=T)%>% mutate(site='peppermint'))
senf_summary <- senf_summary[c(14, 1:13)]
kable(senf_summary, caption = "Sequoia National Forest") %>% kable_styling()

seki_summary <- bind_rows(rwi.stats(cm, cm_ids, prewhiten=T)%>% mutate(site='crescent meadow'),
                        rwi.stats(pr, pr_ids, prewhiten=T)%>% mutate(site='pine ridge'))
seki_summary <- seki_summary[c(14, 1:13)]
kable(seki_summary, caption = "Sequoia National Park") %>% kable_styling()

mamm_summary <- bind_rows(rwi.stats(ic, ic_ids, prewhiten=T)%>% mutate(site='inyo craters'),
                        rwi.stats(sl, sl_ids, prewhiten=T)%>% mutate(site='sherwin lakes'))
mamm_summary <- mamm_summary[c(14, 1:13)]
kable(mamm_summary, caption = "Mammoth Lakes") %>% kable_styling()
```

## SPEI calculation 

Merged PRISM data from Bob and email request for recent years. Calculate SPEI from PET (Thornwaithe) and include this with the long-form PRISM data.

```{r climate data, echo=T, results='hide'}
sitelocations <- read.csv("Raw Data/locations for prism.csv") 
recentyears <- read.csv("Raw Data/prism data/PRISM_ppt_tmin_tmean_tmax_stable_800m_201101_201812_Diez_sites_v2.csv", skip=9, header=T) %>%
  mutate(Elevation_m=Elevation..ft.*0.3048,
         ppt_mm=ppt..inches.*25.4,
         tmean_C=(tmean..degrees.F.-32)*(5/9),
         Date=as.character(Date),
         Name=as.character(Name)) %>%
  dplyr::select(c(Name, Longitude, Latitude, Date, ppt_mm, tmean_C)) %>%
  na.omit() %>%
  filter(Date>=2015)

bobtemp <- read.csv("Raw Data/prism data/tmean189501_201412.csv", header=T) %>%  
  gather(key="Date", value="tmean_C", -c(place, y, x)) %>%
  rename(Name=place, Longitude=x, Latitude=y) %>%
  mutate(Year=str_sub(Date, 2, 5),
         Month=str_sub(Date, 6, 7),
         Date=paste(Year, Month, sep='-'),
         Name=as.character(Name)) %>%
  dplyr::select(Name, Longitude, Latitude, Date, tmean_C) #%>% 
  # filter(Date<2011)

bobppt <- read.csv("Raw Data/prism data/ppt.csv", header=T) %>%
  gather(key="Date", value="ppt_mm", -c(place, y, x)) %>%
  rename(Name=place, Longitude=x, Latitude=y) %>%
  mutate(Year=str_sub(Date, 2, 5),
         Month=str_sub(Date, 6, 7),
         Date=paste(Year, Month, sep='-')) %>%
  dplyr::select(Name, Longitude, Latitude, Date, ppt_mm) #%>%
  # filter(Date<2011)

bobdata <- left_join(bobtemp, bobppt)

allprism <- bind_rows(recentyears, bobdata) %>%
  mutate(Year=as.numeric(str_sub(Date, 1, 4)),
         Month=as.numeric(str_sub(Date, 6,7)))%>%
  arrange(Name, Date)

## SPEI calculation starts here
wide_tmean <- pivot_wider(allprism, id_cols=-c(Latitude, Longitude, ppt_mm), 
                          names_from = Name, values_from = tmean_C) %>%
  arrange(Year, Month)
wide_ppt <- pivot_wider(allprism, id_cols=-c(Latitude, Longitude, tmean_C), 
                        names_from = Name, values_from = ppt_mm)%>%
  arrange(Year, Month)

## PET and Balance calculations (need to be in matrix form for multi-site calculations)
pet <- as.data.frame(thornthwaite(wide_tmean[,4:27], sitelocations$Y)) 
bal <- wide_ppt[,4:27] - pet

ppt_ts  <- ts(wide_ppt[,4:27], end=c(2018, 12), frequency=12) 
prism_ts <- ts(bal, end=c(2018, 12), frequency=12) ## time series for following calculations

spei1 <- spei(prism_ts, 1) ## one-month SPEI
spei1_df <- as.data.frame(spei1$fitted) %>%
  mutate(Year=wide_tmean$Year,
         Month=wide_tmean$Month) %>%
  gather(key="Name", value="spei1",-c(Year, Month))
spei6 <- spei(prism_ts, 6) ## six-month SPEI
spei6_df <- as.data.frame(spei6$fitted) %>%
  mutate(Year=wide_tmean$Year,
         Month=wide_tmean$Month) %>%
  gather(key="Name", value="spei6",-c(Year, Month))
spei12 <- spei(prism_ts, 12) ## twelve-month SPEI
spei12_df <- as.data.frame(spei12$fitted) %>%
  mutate(Year=wide_tmean$Year,
         Month=wide_tmean$Month) %>%
  gather(key="Name", value="spei12",-c(Year, Month))
spei18 <- spei(prism_ts, 18) ## twelve-month SPEI
spei18_df <- as.data.frame(spei18$fitted) %>%
  mutate(Year=wide_tmean$Year,
         Month=wide_tmean$Month) %>%
  gather(key="Name", value="spei18",-c(Year, Month))

spi12 <- spei(ppt_ts, 12) ## twelve-month SPEI
spi12_df <- as.data.frame(spi12$fitted) %>%
  mutate(Year=wide_tmean$Year,
         Month=wide_tmean$Month) %>%
  gather(key="Name", value="spi12",-c(Year, Month))
spi18 <- spei(ppt_ts, 18) ## twelve-month SPEI
spi18_df <- as.data.frame(spi18$fitted) %>%
  mutate(Year=wide_tmean$Year,
         Month=wide_tmean$Month) %>%
  gather(key="Name", value="spi18",-c(Year, Month))

allspei <- left_join(spei1_df, spei6_df) %>% left_join(spei12_df)%>% left_join(spei18_df)%>% left_join(spi12_df)%>% left_join(spi18_df)%>%
  mutate(Site=str_sub(Name, 1,2),
         Neighborhood=as.numeric(str_sub(Name, 3,3))) %>% 
  left_join(sites, by=c("Site"="Site2"))

sept_spei <- filter(allspei, Month==9)

allprism2 <- allprism%>%
  mutate(hydroyear = ifelse(Month>9, Year+1, Year)) %>% 
  group_by(Name, hydroyear) %>% 
  mutate(total_ppt_mm = sum(ppt_mm),
         mean_temp_C = mean(tmean_C)) %>% 
  filter(Month==9) %>% 
  left_join(sept_spei) %>% 
  dplyr::select(-c(ppt_mm, tmean_C))

allprism3 <- allprism%>%
  mutate(hydroyear = ifelse(Month>9, Year+1, Year)) %>% 
  group_by(Name, hydroyear) %>% 
  left_join(allspei) 

# allprism2 <- pivot_wider(allprism, -c(Date, Year, ppt_mm, tmean_C), names_from=Month, values_from=c(spei1, spei6, spei12, spei18, spi12, spi18)) %>%
#   mutate(Site=str_sub(Name, 1,2), Neighborhood=str_sub(Name,3,3))
write.csv(allprism2, "Processed Data/precip_temp_spei.csv", row.names = F)
write.csv(allprism3, "Processed Data/monthly_precip_temp_spei.csv", row.names = F)
```

## Field data
This includes location, DBH, site characteristics, and hegyi index calculation.  Still a handful of id problems, verify field and gps data.

```{r manip, echo=T, results='hide'}
# dia_data2 <- dia_data %>% 
#   dplyr::select(-Region) %>% 
#   left_join(sites2) %>% 
#   mutate(Species=tolower(paste(str_sub(Species, 1,1), str_sub(Species, 3,3), sep="")),
#          Species=ifelse(Species=="?", "xx",
#                          ifelse(Species=="x", "xx",
#                                 ifelse(Species=="a?", "ax", Species))),
#          tree.uniqueID = tolower(paste(paste(siteno, Species, Neighborhood, sep=""), ID, sep="_")),
#          DBHstem1=ifelse(!is.na(DBHT3), DBHT3, 
#                        ifelse(!is.na(DBHT2), DBHT2, DBH)),
#          BA1 = .00007854*DBHstem1^2,
#          BA2 = .00007854*DBHstem2^2,
#          BA3 = .00007854*DBHstem3^2) %>% 
#   mutate(totDBH=rowSums(dplyr::select(.,"DBHstem1", "DBHstem2", "DBHstem3"), na.rm=T),
#          totBA=rowSums(dplyr::select(.,"BA1", "BA2", "BA3"), na.rm=T),
#          totDBH=ifelse(totDBH==0, NA, totDBH),
#          totBA=ifelse(totBA==0, NA, totBA)) %>% 
#   dplyr::select(c(tree.uniqueID, Site, Neighborhood, Species, totDBH, totBA))

library(lubridate)

dia_data2 <- dia_data_new %>%
  dplyr::select(-Region, -Site) %>%
  left_join(sites2) %>%
  mutate_at(c("DBH", "DBHstem2", "DBHstem3", "DBHT2", "DBHT3", "Neighborhood"), as.numeric) %>%
  mutate_at(vars(starts_with("Status")), as.numeric) %>% 
  mutate_at(vars(starts_with("Date")), ymd) %>% 
  mutate(
         DBHstem1=ifelse(!is.na(DBHT3), DBHT3,
                       ifelse(!is.na(DBHT2), DBHT2, DBH)),
         BA1 = .00007854*DBHstem1^2,
         BA2 = .00007854*DBHstem2^2,
         BA3 = .00007854*DBHstem3^2) %>%
  mutate(totDBH=rowSums(dplyr::select(.,"DBHstem1", "DBHstem2", "DBHstem3"), na.rm=T),
         totBA=rowSums(dplyr::select(.,"BA1", "BA2", "BA3"), na.rm=T),
         totDBH=ifelse(totDBH==0, NA, totDBH),
         totBA=ifelse(totBA==0, NA, totBA))


write.csv(dia_data2, "Processed Data/field_data.csv", row.names = F)

dia_data2 <- dia_data2 %>%
  dplyr::select(c(tree.uniqueID, Site, Neighborhood, Species, totDBH, totBA))

raw_loc <- read.csv("Raw data/all_loc.csv") 

test <- raw_loc %>% 
  filter(tree.uniqueID %in% raw_loc$tree.uniqueID[duplicated(raw_loc$tree.uniqueID)==T])

loc_data <- raw_loc %>% 
  inner_join(dplyr::select(dia_data2, c(tree.uniqueID, Site, Neighborhood))) %>% ## there are some naming errors
  group_by(tree.uniqueID) %>% summarise_all(first)
# loc_data <- loc_data %>%
#   filter(Site=="Black Mountain") %>%
#   filter(Neighborhood==3)

loc_grid <- loc_data %>% 
  rename(Site.x=Site, Neighborhood.x=Neighborhood, tree.uniqueID.x=tree.uniqueID, X.x=X, Y.x=Y) %>% 
  expand_grid(dplyr::select(loc_data, c(Site, Neighborhood, tree.uniqueID, X, Y))) %>% 
  filter(Site.x==Site, Neighborhood.x==Neighborhood) %>% 
  dplyr::select(-c(Site.x, Neighborhood.x)) 

library(geosphere)
loc_grid2 <- loc_grid %>% 
  mutate(dist_m=distGeo(dplyr::select(loc_grid, c(X.x, Y.x)), dplyr::select(loc_grid, c(X, Y))),
         bearing_deg=round(bearing(dplyr::select(loc_grid, c(X.x, Y.x)), dplyr::select(loc_grid, c(X, Y))), 2))

xy <- dplyr::select(loc_data, c(X, Y))
pair <- data.frame(distm(xy, fun=distGeo)) %>% ## calculates distance in m based on WGS84 elipsoid
  mutate(tree.uniqueID=loc_data$tree.uniqueID)
names(pair) <- c(loc_data$tree.uniqueID, "focalID")

site_area <- loc_data %>% 
  group_by(Site, Neighborhood) %>% 
  summarise(area=NA)


for(i in unique(loc_data$Site)) {
  for(j in 1:3) {
    data <- filter(loc_data, Site==i, Neighborhood==j) %>% 
      dplyr::select(X,Y)
    ch <- chull(data)
    coords <- data[c(ch, ch[1]), ] 
    site_area$area[which(site_area$Site==i&site_area$Neighborhood==j)] <- areaPolygon(coords)
  }
}
# 
# plot(data)
# lines(coords, col="red")

write.csv(site_area, "Processed Data/site_area.csv", row.names = F)

dist_focal <- pair %>% 
  pivot_longer(cols=-focalID, names_to="periphID", values_to="dist") %>% 
  filter(dist!=0) %>% 
  filter(focalID %in% unique(allrings$tree.uniqueID)) %>% 
  filter(periphID %in% unique(allrings$tree.uniqueID)) %>% 
  left_join(dplyr::select(dia_data2, c("focalID"="tree.uniqueID", Species, totDBH, Site, Neighborhood))) %>% 
  rename(focalsp=Species, DBH=totDBH, Site.x=Site, Neighborhood.x=Neighborhood) %>% 
  left_join(dplyr::select(dia_data2, c("periphID"="tree.uniqueID", Species, totBA, Site, Neighborhood))) %>% 
  rename(periphsp=Species) %>% 
  filter(Site==Site.x, Neighborhood==Neighborhood.x) %>% 
  dplyr::select(-c(Neighborhood.x, Site.x, Neighborhood, Site))

dist_long <- pair %>% 
  pivot_longer(cols=-focalID, names_to="periphID", values_to="dist") %>% 
  filter(dist!=0, dist<=20) %>% 
  filter(focalID %in% unique(allrings$tree.uniqueID)) %>% 
  left_join(dplyr::select(dia_data2, c("focalID"="tree.uniqueID", Species, totDBH, totBA))) %>% 
  rename(focalsp=Species, focalDBH=totDBH, focalBA=totBA) %>% 
  left_join(dplyr::select(dia_data2, c("periphID"="tree.uniqueID", Species, totDBH, totBA))) %>% 
  rename(periphsp=Species) %>% 
  mutate(con_het=ifelse(focalsp==periphsp, "con", "het"))

hegyi <- dist_long %>%
  na.omit() %>%
  group_by(focalID, focalDBH, focalBA) %>%
  mutate(ratio=(totDBH/focalDBH)/dist,
         comp_stems=length(periphID),
         compBA=sum(totBA)) %>%
  summarise(hegyi=sum(ratio, na.rm=T),
            BA = sum(compBA, na.rm=T)) 

# plot(hegyi_old$BA, hegyi$BA)
# hegyi_old <- hegyi
write.csv(hegyi, "Processed Data/comp_density.csv", row.names=F)

rings_field <- left_join(allrings, hegyi, by=c("tree.uniqueID"="focalID"))
```

## Resistance, Recovery, Resilience (RRR) data
Calculating mean for 3 years before and 3 years after, as well as resistance, recovery, and resilience. 

```{r 3years, echo=F, results='hide', eval=F}
rings_field$prev3 <- NA
rings_field$next3 <- NA
for(j in unique(rings_field$tree.uniqueID)) {
  for(i in (min(rings_field$Year[which(rings_field$tree.uniqueID==j)])+3):(max(rings_field$Year[which(rings_field$tree.uniqueID==j)])-3)) {
    rings_field$prev3[which(rings_field$tree.uniqueID==j&rings_field$Year==i)] <- mean(c(rings_field$BAI[which(rings_field$tree.uniqueID==j&rings_field$Year==i-3)],
                                                          rings_field$BAI[which(rings_field$tree.uniqueID==j&rings_field$Year==i-2)],
                                                          rings_field$BAI[which(rings_field$tree.uniqueID==j&rings_field$Year==i-1)]), na.rm=T)
    rings_field$next3[which(rings_field$tree.uniqueID==j&rings_field$Year==i)] <- mean(c(rings_field$BAI[which(rings_field$tree.uniqueID==j&rings_field$Year==i+3)],
                                                          rings_field$BAI[which(rings_field$tree.uniqueID==j&rings_field$Year==i+2)],
                                                          rings_field$BAI[which(rings_field$tree.uniqueID==j&rings_field$Year==i+1)]), na.rm=T)
  }
}

allprism2$spei_prev3 <- NA
allprism2$spei_next3 <- NA
for(j in unique(allprism2$Site)) {
  for(k in unique(allprism2$Neighborhood)) {
    for(i in 1899:2018) {
      allprism2$spei_prev3[which(allprism2$Site==j&allprism2$Neighborhood==k&allprism2$Year==i)] <- mean(c(allprism2$spei12[which(allprism2$Site==j&allprism2$Neighborhood==k&allprism2$Year==i-3)],
                                                          allprism2$spei12[which(allprism2$Site==j&allprism2$Neighborhood==k&allprism2$Year==i-2)],
                                                          allprism2$spei12[which(allprism2$Site==j&allprism2$Neighborhood==k&allprism2$Year==i-1)]), na.rm=T)
    allprism2$spei_next3[which(allprism2$Site==j&allprism2$Neighborhood==k&allprism2$Year==i)] <- mean(c(allprism2$spei12[which(allprism2$Site==j&allprism2$Neighborhood==k&allprism2$Year==i+3)],
                                                          allprism2$spei12[which(allprism2$Site==j&allprism2$Neighborhood==k&allprism2$Year==i+2)],
                                                          allprism2$spei12[which(allprism2$Site==j&allprism2$Neighborhood==k&allprism2$Year==i+1)]), na.rm=T)
    }
  }
}
```

#### Drought cutoff can be changed here:

Refer to the "DroughtSelection" shiny for more information. I used the bottom 10th percentile for drought years and only drought years with associated growth reduction.

```{r cutoff, echo=T, results='hide'}
climate <- allprism2 %>% 
  group_by(Site, Neighborhood) %>% 
  dplyr::select(-c(siteno, Region, hilo, spei18, spi18)) %>% 
  filter(!is.na(spei12)) %>% 
  mutate(cutoff=quantile(spei12, .10, na.rm=T), ### Here is the drought cutoff
         droughtyear=ifelse(spei12<cutoff, "yes","no")) %>% ungroup()
```


```{r RRR, echo=F, eval=F}
rrr_data <- left_join(rings_field, climate, by=c("Site", "Neighborhood", "Year")) %>% 
  filter(droughtyear=="yes") %>% 
  mutate(reduction=ifelse(BAI<prev3, "yes", "no"))

length(which(rrr_data$reduction=="yes"))
length(which(rrr_data$reduction=="no"))

rrr_data <- rrr_data %>% 
  mutate(resistance = BAI/prev3,
         recovery = next3/BAI,
         resilience = next3/prev3,
         rel_res = resilience-resistance,
         drought_fx3 = (prev3-BAI)/BAI)


write.csv(rrr_data, "Processed Data/RRR_data.csv", row.names = F)
```

## Drought effect data
Drought effects sensu Kannenberg et al 2019

```{r drought_fx, echo=T, results='hide'}
dfx_data <- inner_join(rings_field, climate, by=c("Site", "Neighborhood", "Year")) 

dfx_data$post1 <- NA
dfx_data$post2 <- NA
dfx_data$spei_post1 <- NA
dfx_data$spei_post2 <- NA

for(j in unique(dfx_data$tree.uniqueID)) {
  for(i in (min(dfx_data$Year[which(dfx_data$tree.uniqueID==j)])):(max(dfx_data$Year[which(dfx_data$tree.uniqueID==j)])-1)) {
    dfx_data$post1[which(dfx_data$tree.uniqueID==j&dfx_data$Year==i)] <- dfx_data$spline_growth[which(dfx_data$tree.uniqueID==j&dfx_data$Year==i+1)]
    dfx_data$spei_post1[which(dfx_data$tree.uniqueID==j&dfx_data$Year==i)] <- dfx_data$spei12[which(dfx_data$tree.uniqueID==j&dfx_data$Year==i+1)]
  }
}

for(j in unique(dfx_data$tree.uniqueID)) {
  for(i in (min(dfx_data$Year[which(dfx_data$tree.uniqueID==j)])):(max(dfx_data$Year[which(dfx_data$tree.uniqueID==j)])-2)) {
    dfx_data$post2[which(dfx_data$tree.uniqueID==j&dfx_data$Year==i)] <- dfx_data$spline_growth[which(dfx_data$tree.uniqueID==j&dfx_data$Year==i+2)]
    dfx_data$spei_post2[which(dfx_data$tree.uniqueID==j&dfx_data$Year==i)] <- dfx_data$spei12[which(dfx_data$tree.uniqueID==j&dfx_data$Year==i+2)]
  }
}

dfx_data <- mutate(dfx_data,
                   Year_post1 = Year+1,
                   Year_post2 = Year+2) %>% 
  rename(DBH=focalDBH)

dfx_summary <- dfx_data %>% 
  group_by(Region, Site, hilo, Species, Neighborhood, tree.uniqueID, droughtyear, hegyi, BA, DBH) %>% 
  summarise(growth=mean(spline_growth, na.rm=T),#fss_growth
            post1=mean(post1, na.rm=T),
            post2=mean(post1, na.rm=T)) %>% 
  pivot_wider(names_from="droughtyear", values_from=c("growth", "post1","post2")) %>% 
  dplyr::select(-c(post1_no, post2_no)) %>% 
  rename(no=growth_no, yes=growth_yes, post1=post1_yes, post2=post2_yes) %>% 
  mutate(diff=no-yes,
         drought_fx=(no-yes)/no,
         post1_fx=(no-post1)/no,
         post2_fx=(no-post2/no))

dfx_yearly <- dfx_data %>% 
  group_by(tree.uniqueID, droughtyear) %>% 
  mutate(avg_growth=mean(spline_growth, na.rm=T), #fss_growth
         diff=avg_growth-spline_growth, #fss_growth
         reduction=ifelse(diff>0, "yes", "no"),
         drought_fx=(avg_growth-spline_growth)/avg_growth, #fss_growth
         post1_fx=(avg_growth-post1)/avg_growth,
         post2_fx=(avg_growth-post2)/avg_growth)

dfx_allyears <- dfx_yearly 

dfx_yearly <- dfx_yearly %>% 
  filter(droughtyear=="yes")

write.csv(dfx_summary, "Processed Data/dfx_summary.csv")
write.csv(dfx_yearly, "Processed Data/dfx_yearly.csv")
write.csv(dfx_allyears, "Processed Data/dfx_allyears.csv")

```

Recovery/legacy is calculated as
$$
Recovery = RWI_{actual} - RWI_{pred}
$$

To calculate $RWI_{pred}$ I modeled RWI as a function Species and SPEI I used the 'brms' package in R. For simplicity I denoted here site-specific regression coefficients, but they were modeled as elevation nested within region.
$$
RWI_{pred_{i}} = Normal(\mu_i, \sigma_i)
$$

$$
\mu_i = \alpha_{neighborhood_i} + \alpha_{treeID_i} + \beta_{1 neighborhood_i}Species_i + \beta_{2 neighborhood_i}SPEI_{i} + \beta_{3 neighborhood_i}Species_i\times SPEI_i
$$

$$
\left[\begin{array}
{rrr}
\alpha_{neighborhood} \\
\beta_{1neighborhood} \\
\beta_{2neighborhood} \\
\beta_{3neighborhood}
\end{array}\right] 

=

MVNormal\left(\left[\begin{array}
{rrr}
\alpha \\
\beta_1 \\
\beta_2 \\
\beta_3 \\
\end{array}\right], \mathbf{S} \right)
$$
I think???? where **S** is the covariance matrix.  Can report priors here.
```{r legacy, echo=T, results='hide'}
# dfx_yearly <- read.csv("Processed Data/dfx_yearly.csv")

mod <- readRDS("Brms models/spline_growth models/spline_growth11_alldata.rds")

dfx_data <- filter(dfx_yearly, zeroRing=="no", Species!="AM") %>% 
  mutate(unique.nb = paste0(Site, Neighborhood, sep="")) %>% 
  ungroup()
dfx_data$Species[str_which(dfx_data$Species, "PP")] <- "PJ" 
# dfx_data$Species <- droplevels(dfx_data$Species)

for(i in unique(dfx_data$tree.uniqueID)){ # unique(dfx_data$tree.uniqueID)
  print(i)
  for(j in unique(dfx_data$Year[which(dfx_data$tree.uniqueID==i)])) {
    newdata <- filter(dfx_data, tree.uniqueID==i&Year==j) %>%
      dplyr::select(c(spei_post1, Species, unique.nb, tree.uniqueID, post1)) %>%
      rename(spei12=spei_post1)

    pred_grow <- fitted(mod, newdata=newdata, summary=FALSE, nsamples = 1000, allow_new_levels=TRUE)
    newy_df <- data.frame(pred_grow) %>%
      mutate(tree.uniqueID=i,
             Year=j,
             post1=newdata$post1,
             diff=post1-pred_grow)
    if(i==first(unique(dfx_data$tree.uniqueID))&j==min(dfx_data$Year[which(dfx_data$tree.uniqueID==i)])) {
      big_df <- newy_df
    } else {big_df <- bind_rows(big_df, newy_df)}
  }
}

# for(i in unique(dfx_data$tree.uniqueID)){ # unique(dfx_data$tree.uniqueID)
#   print(i)
#   for(j in unique(dfx_data$Year[which(dfx_data$tree.uniqueID==i)])) {
#     newdata <- filter(dfx_data, tree.uniqueID==i&Year==j) %>% 
#       dplyr::select(c(spei_post2, Species, unique.nb, tree.uniqueID, post2)) %>% 
#       rename(spei12=spei_post2)
#     
#     pred_grow <- fitted(mod, newdata=newdata, summary=FALSE, nsamples = 500)
#     newy_df <- data.frame(pred_grow) %>% 
#       mutate(tree.uniqueID=i,
#              Year=j,
#              post2=newdata$post2,
#              diff=post2-pred_grow)
#     if(i==first(unique(dfx_data$tree.uniqueID))&j==min(dfx_data$Year[which(dfx_data$tree.uniqueID==i)])) {
#       big_df2 <- newy_df
#     } else {big_df2 <- bind_rows(big_df2, newy_df)}
#   }
# }

legacy <- big_df %>%
  group_by(tree.uniqueID, Year) %>%
  summarise(pred_mean = mean(pred_grow, na.rm=T),
            pred_50 = quantile(pred_grow, .5, na.rm=T),
            pred_025 = quantile(pred_grow, .025, na.rm=T),
            pred_975 = quantile(pred_grow, .975, na.rm=T),
            leg1_mean = mean(diff, na.rm=T),
            leg1_50 = quantile(diff, .5, na.rm=T),
            leg1_025 = quantile(diff, .025, na.rm=T),
            leg1_975 = quantile(diff, .975, na.rm=T)) %>%
  left_join(dplyr::select(dfx_data, c(Year, Site, Species, Neighborhood, Region, hilo, tree.uniqueID, hegyi, BA, DBH, reduction,
                               spei12, post1, post2, spei_post1, spei_post2, Year_post1, Year_post2, drought_fx)))


# legacy2 <- big_df2 %>% 
#   group_by(tree.uniqueID, Year) %>% 
#   summarise(pred_mean = mean(pred_grow, na.rm=T),
#             pred_50 = quantile(pred_grow, .5, na.rm=T),
#             pred_025 = quantile(pred_grow, .025, na.rm=T),
#             pred_975 = quantile(pred_grow, .975, na.rm=T),
#             leg2_mean = mean(diff, na.rm=T),
#             leg2_50 = quantile(diff, .5, na.rm=T),
#             leg2_025 = quantile(diff, .025, na.rm=T),
#             leg2_975 = quantile(diff, .975, na.rm=T)) %>% 
#   dplyr::select(-c(pred_mean, pred_50, pred_025, pred_975)) %>% 
#   left_join(legacy)

legacy$X <- NULL
l_leg <- pivot_longer(legacy, c(leg1_mean, leg1_50, leg1_025, leg1_975
                                ), names_to = c("event","qi"), names_sep="_", #,leg2_mean, leg2_50, leg2_025, leg2_975
                      values_to = "value") %>% 
  dplyr::select(-c(spei_post1, Year_post1, Year_post2)) %>% 
  pivot_wider(names_from="qi", values_from="value") %>% 
  rename(legacy_mean=mean, legacy_50=`50`, legacy_025=`025`, legacy_975=`975`) %>% 
  mutate(event=ifelse(event=="leg1", "post1", "post2"))

l_leg2 <- dplyr::select(legacy, c(tree.uniqueID, Year, spei_post1, spei_post2, Year_post1, Year_post2)) %>% 
  pivot_longer(c(spei_post1, spei_post2, Year_post1, Year_post2), names_to = c("dummy", "event"), 
               names_sep="_", values_to="value") %>% 
  mutate(dummy=ifelse(dummy=="spei", "spei_post", "year_post")) %>% 
  pivot_wider(names_from="dummy", values_from="value") %>% 
  left_join(l_leg) %>% 
  ungroup() %>% 
  mutate(Region1 =ifelse(Region=="San Jac", "lat1", ifelse(Region=="SENF", "lat2", ifelse(Region=="SEKI", "lat3", "lat4"))),
         recovery=legacy_mean)

l_leg1 <- filter(l_leg2, event=="post1")

# write.csv(l_leg2, "Processed Data/legacy_long.csv", row.names = F)
write.csv(l_leg1, "Processed Data/legacy_1post.csv", row.names = F)
```

```{r legacy_ar, echo=F, eval=F}
# dfx_yearly <- read.csv("Processed Data/dfx_yearly.csv")

mod <- readRDS("Brms models/spline_growth models/spline_growth12_alldata.rds")

dfx_data <- filter(dfx_yearly, zeroRing=="no", Species!="AM") %>% 
   mutate(unique.nb = paste0(Site, Neighborhood, sep=""))
dfx_data$Species[str_which(dfx_data$Species, "PP")] <- "PJ" 
# dfx_data$Species <- droplevels(dfx_data$Species)

for(i in unique(dfx_data$tree.uniqueID)){ # unique(dfx_data$tree.uniqueID)
  print(i)
  for(j in unique(dfx_data$Year[which(dfx_data$tree.uniqueID==i)])) {
    newdata <- filter(dfx_data, tree.uniqueID==i&Year==j) %>%
      dplyr::select(c(spei_post1, Species, Year, unique.nb, tree.uniqueID, post1)) %>%
      rename(spei12=spei_post1)

    pred_grow <- fitted(mod, newdata=newdata, summary=FALSE, allow_new_levels=T, nsamples = 500)
    newy_df <- data.frame(pred_grow) %>%
      mutate(tree.uniqueID=i,
             Year=j,
             post1=newdata$post1,
             diff=post1-pred_grow)
    if(i==first(unique(dfx_data$tree.uniqueID))&j==min(dfx_data$Year[which(dfx_data$tree.uniqueID==i)])) {
      big_df <- newy_df
    } else {big_df <- bind_rows(big_df, newy_df)}
  }
}

legacy <- big_df %>% 
  group_by(tree.uniqueID, Year) %>% 
  summarise(pred_mean = mean(pred_grow, na.rm=T),
            pred_50 = quantile(pred_grow, .5, na.rm=T),
            pred_025 = quantile(pred_grow, .025, na.rm=T),
            pred_975 = quantile(pred_grow, .975, na.rm=T),
            leg1_mean = mean(diff, na.rm=T),
            leg1_50 = quantile(diff, .5, na.rm=T),
            leg1_025 = quantile(diff, .025, na.rm=T),
            leg1_975 = quantile(diff, .975, na.rm=T)) %>% 
  left_join(dplyr::select(dfx_data, c(Year, Site, Species, Neighborhood, Region, hilo, tree.uniqueID, hegyi, DBH, reduction,
                               spei12, post1, post2, spei_post1, spei_post2, Year_post1, Year_post2, drought_fx)))


legacy$X <- NULL
l_leg <- pivot_longer(legacy, c(leg1_mean, leg1_50, leg1_025, leg1_975
                                ), names_to = c("event","qi"), names_sep="_", #,leg2_mean, leg2_50, leg2_025, leg2_975
                      values_to = "value") %>% 
  dplyr::select(-c(spei_post1, spei_post2, Year_post1, Year_post2)) %>% 
  pivot_wider(names_from="qi", values_from="value") %>% 
  rename(legacy_mean=mean, legacy_50=`50`, legacy_025=`025`, legacy_975=`975`) %>% 
  mutate(event=ifelse(event=="leg1", "post1", "post2"))

l_leg2 <- dplyr::select(legacy, c(tree.uniqueID, Year, spei_post1, spei_post2, Year_post1, Year_post2)) %>% 
  pivot_longer(c(spei_post1, spei_post2, Year_post1, Year_post2), names_to = c("dummy", "event"), 
               names_sep="_", values_to="value") %>% 
  mutate(dummy=ifelse(dummy=="spei", "spei_post", "year_post")) %>% 
  pivot_wider(names_from="dummy", values_from="value") %>% 
  left_join(l_leg) %>% 
  group_by(event) %>% 
  mutate(scaled_leg=(legacy_mean-min(legacy_mean, na.rm=T))/(max(legacy_mean, na.rm=T)-min(legacy_mean, na.rm=T)),
         scaled_dfx=(drought_fx-min(drought_fx, na.rm=T))/(max(drought_fx, na.rm=T)-min(drought_fx, na.rm=T)),
         scaled_spei=(1-(spei12-min(spei12, na.rm=T))/(max(spei12, na.rm=T)-min(spei12, na.rm=T))),
         scaled_spei=ifelse(scaled_spei==0, 0.0001, scaled_spei),
         Region1 =ifelse(Region=="San Jac", "lat1", ifelse(Region=="SENF", "lat2", ifelse(Region=="SEKI", "lat3", "lat4")))) %>% 
  ungroup() %>% 
  mutate(rel_dfx=scaled_dfx/scaled_spei,
         rel_leg=scaled_leg/scaled_dfx,
         recovery=legacy_mean)

l_leg1 <- filter(l_leg2, event=="post1")

# write.csv(l_leg2, "Processed Data/legacy_long.csv", row.names = F)
write.csv(l_leg1, "Processed Data/legacy_1post_with ar.csv", row.names = F)

```
